A Theory for Data Migration of Information Systems

### 2022-10-22 2022 Migration/articleMigration.tex
minor changes in abstract and introduction

\begin{abstract}
   The Ampersand project has provided the theory and tools to generate semantic information systems from an algebraic specification.
   However, information systems in practice may change repeatedly after their maiden deployment.
   Changes that affect the data model typically result in a data migration.
   In such cases, simply regenerating the system is not enough because it would reset the database to its initial state (deleting all data gathered so far).
   Data migration aims at preserving the data as much as possible.
   Typically, migration involves transferring the old data to the new system while preserving the semantics as much as possible.

@@

\section{Introduction}
\label{sct:Introduction}
   Our research focuses on generating information systems incrementally,
   especially on the data migration aspect of it.
   The reason that we believe in generating information systems is to prevent human errors as early as possible in the build chain:
   A mistake not made cannot propagate into production.
   For this purpose, we use an existing compiler, Ampersand~\cite{Joosten-JLAMP2018},
   that generates an information system from an Ampersand script.
   An Ampersand script contains just enough information to generate a complete information system,
   which means that a classical database schema (i.e. data structure plus semantics) can be extracted from the Ampersand script.
   For the purpose of this paper we can equate an Ampersand script with the scheme of the generated information system.

   Incremental generation of information systems poses specific requirements to data migration.
   For instance, if a data migration is not for incremental generation,
   e.g.~\cite{Gholami2016,Bisbal1999},
   migration engineers may want to preserve the functionality to avoid introducing new errors in an otherwise error-prone process.
   For incremental generation, however, preserving the functionality is not an option.
   To change the functionality is the very purpose and the data migration must behave accordingly.
   So, we must make sure to use methods for data migration that are specifically meant for incrementally generating information systems.

@@

   Much like~\cite{Spivak2012}, our work assumes that the semantics of the as-is dataset is encoded in its schema,
   so we can be explicit about ``preserving the meaning as much as possible''.
   The schema $\schema_{\infsys}$ of an information system $\infsys$ is basically a set of constraints on its dataset $\dataset$.
   (We will denote the schema just as $\schema$ if the information system is obvious or irrelevant to know.)
   In contrast with~\cite{Spivak2012}, our approach does not work with functions (functors)
   but with relations in a way that bears resemblance with~\cite{Zielinski2013}.

>>>>

\begin{abstract}
   The Ampersand project has provided the theory and tools to generate semantic information systems from an algebraic specification.
   However, information systems in practice may change repeatedly after their maiden deployment.
   Changes that affect the data model typically result in a data migration.
   In such cases, simply regenerating the system is not enough because it would reset the database to its initial state (deleting all data gathered so far).
   A migration engineer must transfer the old data to the new system while preserving the semantics as much as possible.

@@

\section{Introduction}
\label{sct:Introduction}
   Our research focuses on generating information systems incrementally,
   especially on the data migration aspect of it.
   The reason that we believe in generating information systems is to prevent human errors as early as possible in the build chain:
   A mistake not made cannot propagate into production.
   For this purpose, we use an existing compiler, Ampersand~\cite{Joosten-JLAMP2018},
   that generates an information system from an Ampersand script.
   An Ampersand script contains just enough information to generate a complete information system,
   which means that a classical database schema (i.e. data structure plus semantics) can be extracted from the Ampersand script.
   For the purpose of this paper we can equate an Ampersand script with the schema of the generated information system.

   Incremental development of information systems poses specific requirements to data migration.
   For instance, in data migrations for other purposes,
   e.g.~\cite{Gholami2016,Bisbal1999},
   migration engineers may want to preserve the functionality to avoid introducing new errors in an otherwise error-prone process.
   For incremental development, however, preserving the functionality is not an option.
   To change the functionality is the very purpose of each increment, so the data migration must support just that.
   So, we must make sure to use methods for data migration that are specifically suitable for incremental development of information systems.

@@

   Much like~\cite{Spivak2012}, our work assumes that the semantics of the as-is dataset is encoded in its schema,
   so we can be explicit about ``preserving the meaning as much as possible''.
   The schema $\schema_{\infsys}$ of an information system $\infsys$ is basically a set of constraints on its dataset $\dataset$.
   (We will denote the schema just as $\schema$ if the information system is obvious or irrelevant to know.)
   In contrast with~\cite{Spivak2012}, our approach does not work with functions (functors)
   but with relations in a way that bears resemblance with~\cite{Alloy2006} and~\cite{Zielinski2013}.
